Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-05-14 WholePlotTrajectories.Rmd
web pages

```{r NMDSplans, out.width = '100%', fig.asp = 1.2/2, echo=FALSE, fig.cap="Trajectories of the plots in terms of flora composition (left panels (ref:Bolda) and (ref:Boldc)) and functional composition (right panels (ref:Boldb) and (ref:Boldd)) regarding the 6 leaf and stem functional traits, the maximum allometric height and seed mass class. Plots trajectories are first represented in the two-dimensional space from the NMDS performed for the 30 years after disturbance based on Bray-Curtis dissimilarity measures between successive inventories (Upper panels (ref:Bolda) and (ref:Boldb)). Then the lower panels ((ref:Boldc) and (ref:Boldd)) represent the euclidean distance to initial condition along the 30 sampled years. Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3). The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.", fig.env="figure*"}
load("DB/FunctionalComposition_forGraph");load("DB/TaxoComposition_forGraphs")
load("DB/FunctionalComposition_forGraph");load("DB/TaxoComposition_forGraphs")

@@

The changes observed in plots functional composition went hand to hand with consistent trajectories of the 8 functional and life history traits (Leaf thickness, chlorophyll content, toughness and specific area, wood specific gravity and bark thickness and seed mass and maximum adult height) visualized with plots community weighted means (CWM) (Figure \@ref(fig:CWM)).

@@

### Taxonomic diversity

@@

For undisturbed plots the Richness, Shannon and Simpson diversities remained stable over the 30 monitored years.
After disturbance the richness increased when the disturbance intensity was low, with a maximum increase of 14 botanical genera (plot 3 from treatment 2).
When the disturbance was intense however, plots taxonomic richness followed a unimodal decrease with a return to initial values after intense disturbance.
The taxonomic evenness (Shannon and Simpson diversities) significantly increased irrespective of the disturbance intensity.
The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum reached around 20 years and positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
Only two T3 plots, plots 8 and 12, remained increasing 30 years after disturbance \@ref(fig:DivTaxo)), suggesting a similar be much delayed unimodal trajectory.

### Functional diversity

The functional diversity trajectories of plots based on the 8 leaf, stem and life history traits was examined through the Rao diversity.
For all undisturbed plots the functional diversity remained comparable to the reference values 5 years after disturbance (1989 inventories).
For all disturbed plots the functional trajectories followed a unimodal trajectory with a maximum positively correlated to disturbance intensity ($\rho_{spearman}=0.73$) (Figure \@ref(fig:IDHplot)). 
The plot 7 from treatment 1 that displayed a constantly outlying diversity was removed from the graphical representation for better readability, see appendix for full graph.
Thirty years after disturbance all plots, whenever the initial disturbance intensity, regained diversity values similar to their initial value and to those of control plots.

```{r IDHplot,  echo=FALSE, fig.cap="Upper panels, Trajectories of the Simpson taxonomic diversity (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance, corresponding to the median and 0.025 and 0.975 percentile observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3. Lower panels, Relationship between the initial \\%AGB removed and the values of Simpson (ref:Boldc) and Rao (ref:Boldd) diversities at three times after disturbance. Solid lines colors represent the time, 10 years (yellow), 20 years (orange) and 30 years (brown) after disturbance.", fig.env="figure*"}

load("DB/SimpsonDraw");load("DB/FunctionalTraj_ForGraphs")
load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB")

@@

plotIDH(Data=Sim,AgbLoss=AGBloss)
plotIDH(Data=Rao,AgbLoss=AGBloss)
mtext("initial %AGB lost",side=1,line=2.2,adj=1)
legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
```

>>>>

```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the plots in terms of flora composition (left panels (ref:Bolda) and (ref:Boldc)) and functional composition (right panels (ref:Boldb) and (ref:Boldd)) regarding the 6 leaf and stem functional traits, the maximum allometric height and seed mass class. Plots trajectories are first represented in the two-dimensional space from the NMDS performed for the 30 years after disturbance based on Bray-Curtis dissimilarity measures between successive inventories (Upper panels (ref:Bolda) and (ref:Boldb)). Then the lower panels ((ref:Boldc) and (ref:Boldd)) represent the euclidean distance to initial condition along the 30 sampled years. Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3). The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.", fig.env="figure*"}
load("DB/FunctionalComposition_forGraph");load("DB/TaxoComposition_forGraphs")
load("DB/FunctionalComposition_forGraph");load("DB/TaxoComposition_forGraphs")

@@

The changes observed in plots functional composition went hand to hand with consistent trajectories of the 8 functional and life history traits visualized through plots community weighted means (CWM) (Figure \@ref(fig:CWM)) of the leaves thickness, chlorophyll content, toughness and specific area, wood specific gravity, bark thickness, seed mass and maximum adult height.

@@

### Communities taxonomic and functional diversity

@@

For undisturbed plots the Richness, Shannon and Simpson diversities remained stable over the 30 years monitored.
After low disturbance intensity the richness increased, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2).
After intense disturbance however, the taxonomic richness followed a unimodal decrease with a return to pre-disturbance values.
The taxonomic evenness (Shannon and Simpson diversities) increased after disturbance, whenever its intensity.
The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum value, reached around 20 years, positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
Only two T3 plots (plots 8 and 12) were still increasing 30 years after disturbance \@ref(fig:DivTaxo)), suggesting a similar be much delayed unimodal trajectory.

Communities functional diversity trajectories were examined through the Rao diversity based on the 7 leaf, stem and life history traits (the seed mass was excluded as a qualitative variable).
The plot 7 from treatment 1 that displayed a constantly outlying diversity was removed from the graphical representation for better readability, see figure A3 in appendix for full graph.
The functional diversity of all undisturbed communities remained stable along the 30 years monitored.
For all disturbed communities the functional diversity followed unimodal trajectories with a marked return towards the initial state (5 years after disturbance) with values comparable to those of control plots.

The impact of disturbance was examined specifically at 10, 20 and 30 years after disturbance through the linear correlation between the initial percentage of AGB removed and the value of communities Simpson and Rao diversities (both diversities of order 2).

```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1/1.5,fig.cap="Upper panels, Trajectories of the Simpson taxonomic diversity (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance, corresponding to the median and 0.025 and 0.975 percentile observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3. Lower panels, Relationship between the initial \\%AGB removed and the median values of Simpson (ref:Boldc) and Rao (ref:Boldd) diversities at three times after disturbance. Solid lines colors represent the time, 10 years (yellow), 20 years (orange) and 30 years (brown) after disturbance.", fig.env="figure*"}
load("DB/SimpsonDraw");load("DB/FunctionalTraj_ForGraphs")
load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB")

@@

plotIDH(Data=Sim,AgbLoss=AGBloss)
plotIDH(Data=Rao,AgbLoss=AGBloss)
mtext("initial %AGB lost",side=1,line=2.2,adj=1)
legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")

### 2018-05-14 WholePlotTrajectories.tex
web pages

The changes observed in plots functional composition went hand to hand
with consistent trajectories of the 8 functional and life history traits
(Leaf thickness, chlorophyll content, toughness and specific area, wood
specific gravity and bark thickness and seed mass and maximum adult
height) visualized with plots community weighted means (CWM) (Figure
\ref{fig:CWM}).

@@

\subsubsection{Taxonomic diversity}\label{taxonomic-diversity}

@@

For undisturbed plots the Richness, Shannon and Simpson diversities
remained stable over the 30 monitored years. After disturbance the
richness increased when the disturbance intensity was low, with a
maximum increase of 14 botanical genera (plot 3 from treatment 2). When
the disturbance was intense however, plots taxonomic richness followed a
unimodal decrease with a return to initial values after intense
disturbance. The taxonomic evenness (Shannon and Simpson diversities)
significantly increased irrespective of the disturbance intensity. The
evenness followed a unimodal trajectory with a just beginning return
towards initial values and a maximum reached around 20 years and
positively correlated to the disturbance intensity
(\(\rho_{spearman}^{Shannon}=0.86\), and
\(\rho_{spearman}^{Simpson}=0.89\)). Only two T3 plots, plots 8 and 12,
remained increasing 30 years after disturbance \ref{fig:DivTaxo}),
suggesting a similar be much delayed unimodal trajectory.

\subsubsection{Functional diversity}\label{functional-diversity}

The functional diversity trajectories of plots based on the 8 leaf, stem
and life history traits was examined through the Rao diversity. For all
undisturbed plots the functional diversity remained comparable to the
reference values 5 years after disturbance (1989 inventories). For all
disturbed plots the functional trajectories followed a unimodal
trajectory with a maximum positively correlated to disturbance intensity
(\(\rho_{spearman}=0.73\)) (Figure \ref{fig:IDHplot}). The plot 7 from
treatment 1 that displayed a constantly outlying diversity was removed
from the graphical representation for better readability, see appendix
for full graph. Thirty years after disturbance all plots, whenever the
initial disturbance intensity, regained diversity values similar to
their initial value and to those of control plots.

@@

{\centering \includegraphics{WholePlotTrajectories_files/figure-latex/IDHplot-1}

@@

\caption{Upper panels, Trajectories of the Simpson taxonomic diversity \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance, corresponding to the median and 0.025 and 0.975 percentile observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3. Lower panels, Relationship between the initial \%AGB removed and the values of Simpson \textbf{(c)} and Rao \textbf{(d)} diversities at three times after disturbance. Solid lines colors represent the time, 10 years (yellow), 20 years (orange) and 30 years (brown) after disturbance.}\label{fig:IDHplot}
\end{figure*}

>>>>

The changes observed in plots functional composition went hand to hand
with consistent trajectories of the 8 functional and life history traits
visualized through plots community weighted means (CWM) (Figure
\ref{fig:CWM}) of the leaves thickness, chlorophyll content, toughness
and specific area, wood specific gravity, bark thickness, seed mass and
maximum adult height.

@@

\subsubsection{Communities taxonomic and functional
diversity}\label{communities-taxonomic-and-functional-diversity}

@@

For undisturbed plots the Richness, Shannon and Simpson diversities
remained stable over the 30 years monitored. After low disturbance
intensity the richness increased, reaching a maximum gain of 14
botanical genera (plot 3 from treatment 2). After intense disturbance
however, the taxonomic richness followed a unimodal decrease with a
return to pre-disturbance values. The taxonomic evenness (Shannon and
Simpson diversities) increased after disturbance, whenever its
intensity. The evenness followed a unimodal trajectory with a just
beginning return towards initial values and a maximum value, reached
around 20 years, positively correlated to the disturbance intensity
(\(\rho_{spearman}^{Shannon}=0.86\), and
\(\rho_{spearman}^{Simpson}=0.89\)). Only two T3 plots (plots 8 and 12)
were still increasing 30 years after disturbance \ref{fig:DivTaxo}),
suggesting a similar be much delayed unimodal trajectory.

Communities functional diversity trajectories were examined through the
Rao diversity based on the 7 leaf, stem and life history traits (the
seed mass was excluded as a qualitative variable). The plot 7 from
treatment 1 that displayed a constantly outlying diversity was removed
from the graphical representation for better readability, see figure A3
in appendix for full graph. The functional diversity of all undisturbed
communities remained stable along the 30 years monitored. For all
disturbed communities the functional diversity followed unimodal
trajectories with a marked return towards the initial state (5 years
after disturbance) with values comparable to those of control plots.

The impact of disturbance was examined more accurately at 10, 20 and 30
years after disturbance through the linear correlation between the
initial percentage of AGB removed and the value of communities Simpson
and Rao diversities (both diversities of order 2).

@@

{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/IDHplot-1}

@@

\caption{Upper panels, Trajectories of the Simpson taxonomic diversity \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance, corresponding to the median and 0.025 and 0.975 percentile observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3. Lower panels, Relationship between the initial \%AGB removed and the median values of Simpson \textbf{(c)} and Rao \textbf{(d)} diversities at three times after disturbance. Solid lines colors represent the time, 10 years (yellow), 20 years (orange) and 30 years (brown) after disturbance.}\label{fig:IDHplot}
\end{figure*}
